---
title: "SCORE Criminology"
author: "Shilaan Alzahawi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here) #for a project-based workflow
library(tidyverse) #dplyr, the pipe, ggplot, etc
library(broom) #for tidy results
#load MASS after data construction, due to conflicts with dplyr::select()
options(scipen = 3)
```

**Reproduction analyst(s):** Shilaan Alzahawi  
**SCORE RR ID:** 2w9mo  
**OSF Project:** <https://osf.io/3j7gq/?view_only=b0e33287a13146d1a2d9c4ad938be094>

I did not deviate from the provided materials. All claims were successfully reproduced at the first attempt. To design the analysis, I consulted the following materials: 

- The original paper: Bersani, B. E.; Doherty, E. (2013). When the ties that bind unwind: Examining the enduring and situational processes of change behind the marriage effect. *Criminology, 51*(2), 399. <https://doi.org/10.1111/1745-9125.12008>
- The data posted at [https://www.nlsinfo.org/investigator/pages/search#](https://www.nlsinfo.org/investigator/pages/search#)
- The bushel claims obtained [here](https://osf.io/5xqeh/?view_only=0498fe041e4c4753b4126ac28e9960a8)


## Read data

```{r read-data}
#Data obtained from the NLS here https://www.bls.gov/nls/nlsy97.htm 
NLSY97 <- read.table(here("data/NLSY97.dat"), sep = " ")

#Run helper functions obtained from the NLS
source(here("code/NLSY97-helper-functions.R"))

#Rename variables using Question names instead of Reference Numbers
NLSY97 <- qnames(NLSY97)
```

## Data cleaning and processing 

### Incarceration data

```{r incarceration}
#Get incarceration data
incarceration <- NLSY97 %>% 
  select(contains("INCARC"))

#Calculate the total number of months respondents were incarcerated
months_incarcerated <- rowSums(incarceration, na.rm = TRUE)
```

### Age

```{r age}
#Age at start of survey
age <- NLSY97$`CV_AGE_12/31/96_1997`

#Calculate age for each respondent at each timepoint (156 timepoints in total)
ages <- 1:nrow(NLSY97) %>% 
  #Sequence from age at t=1 to 13 years later
  map(~ seq(age[.x], age[.x] + 13, length.out = 156)) %>% 
  unlist()

#Create dataframe with respondent id and age over time
all_ages <- tibble(
  id = rep(1:nrow(NLSY97), each = 156),
  age = ages
)
```



### Create initial dataset

```{r dataset}
#Create clean dataframe with relevant variables
df <- tibble(
  #respondent id
  id = 1:nrow(NLSY97),
  #gender: 1 = male
  male = ifelse(NLSY97$KEY_SEX_1997 == 2, 0, 1),
  #race: 1 = black
  black = ifelse(NLSY97$KEY_RACE_ETHNICITY_1997 == 1, 1, 0),
  #household size: number of individuals below 18
  household_size = NLSY97$CV_HH_UNDER_18_1997,
  #family structure: 1 = intact
  family_structure = ifelse(NLSY97$CV_YTH_REL_HH_CURRENT_1997 == 1, 1, 0),
  #lack of heat: 1 = home lacked electricity and heat in previous month
  lack_heat = ifelse(NLSY97$`YSAQ-003_1997` == 0, 1, 0),
  #hard times
  hard_times = NLSY97$`PC8-090_1997`,
  #presence of dilapidated buildings: 1 = poorly kept buildings
  dilapidated = ifelse(NLSY97$`YIR-1400_1998` == 3, 1, 0),
  #education of parents
  education_mother = NLSY97$CV_HGC_BIO_MOM_1997,
  education_father = NLSY97$CV_HGC_BIO_DAD_1997,
  #age of first sexual intercourse: 1 = at or before 13 years of age
  age_sex = ifelse(NLSY97$`YSAQ-300_1999` < 14, 1, 0),
  #age of first arrest: 1 = at or before 13 years of age
  age_arrest = ifelse(NLSY97$`YSAQ-441_1997` < 14, 1, 0),
  #age of first weed use: 1 = at or before 13 years of age
  age_weed = ifelse(NLSY97$`YSAQ-370_1997` < 14, 1, 0),
  #age of first cocaine use: 1 = at or before 13 years of age
  age_cocaine = ifelse(NLSY97$`YSAQ-372C_1998` < 14, 1, 0),
  #incarcerated: 1 = at least one episode of incarceration
  incarcerated = ifelse(months_incarcerated > 0, 1, 0)
) %>% 
  rowwise() %>% 
  mutate(
  #disadvantaged = 1 if lack_heat or hard_times or dilapidated buildings
  disadvantage = ifelse(
    sum(lack_heat, hard_times, dilapidated, na.rm = TRUE) > 0, 1, 0
    ),
  #mean parental education
  parental_education = mean(c(education_mother, education_father), na.rm = TRUE),
  #drug use = 1 if used weed or cocaine before 14 years of age
  age_drug = ifelse(sum(age_weed, age_cocaine, na.rm = TRUE) > 0, 1, 0),
  #at risk = sum of age_sex, age_drug, age_arrest
  at_risk = sum(age_sex, age_drug, age_arrest, na.rm = TRUE)
  ) %>% 
  #keep relevant variables
  select(
    id:family_structure, 
    disadvantage:parental_education, 
    at_risk, 
    incarcerated
    )
```

### Arrest status

Arrest status is collected every month, with the following answer options:

- 0: R not arrested in this month and not arrested in a previous month
- 1 to 98: Number of times R arrested in this month
- 99: R arrested previously but not in this month

The authors recode this to a binary, monthly arrest status variable: nonevent months are coded 0, whereas months with at least one documented arrest are coded 1. 

```{r arrest}
arrest <- NLSY97 %>% 
  #select relevant columns
  select(contains("ARREST")) %>% 
  #create participant id
  mutate(
    id = 1:nrow(NLSY97)
  ) %>% 
  #make data long: each time point becomes an individual row
  pivot_longer(
    cols = contains("ARREST"),
    names_to = "time",
    names_pattern = "ARREST_STATUS_?(.*)_XRND",
    values_to = "arrest"
  ) %>% 
  group_by(id) %>% 
  #calculate the total number of arrests for each respondent
  mutate(
    sum_arrests = sum(arrest, na.rm = TRUE)
  ) %>% 
  ungroup() %>% 
  #keep only respondents who were arrested at least once
  filter(sum_arrests > 0) %>% 
  #recode arrest to binary indicator variable
  mutate(
    arrest = ifelse(arrest %in% c(0,99), 0, ifelse(is.na(arrest), NA, 1))
  ) %>% 
  select(!sum_arrests)

#data quality checks
#for each respondent, we should have 156 rows (13 years * 12 months) 
#number_of_timepoints <- ncol(NLSY97 %>% select(contains("ARREST")))
#number_of_respondents <- length(unique(arrest$id)) #in paper: 2,959 (Table 2)
#nrow(arrest) == number_of_respondents * number_of_timepoints #TRUE
```

### Relationship status

Relationship status is collected every month, with the following answer options:

- 0 Never Married, Not Cohabitating
- 1 Never Married, Cohabiting
- 2 Married
- 3 Legally Separated
- 4 Divorced
- 5 Widowed

The authors create separate, mutually exclusive relationship indicators:

- Single
- Married
- Legally separated
- Divorced

It is unclear what the authors do with widowers.

```{r marital}
marital <- NLSY97 %>% 
  #select relevant columns
  select(contains("MAR")) %>% 
  #create participant id
  mutate(
    id = 1:nrow(NLSY97)
  ) %>% 
  #make data long: each time point becomes an individual row
  pivot_longer(
    cols = contains("MAR"),
    names_to = "time",
    names_pattern = "MAR_STATUS_?(.*)_XRND",
    values_to = "marital_status"
  ) %>%
  #create indicator variables for relationship status
  mutate(
    #note that the paper does not disclose what to do with widowers
    married = ifelse(marital_status == 2, 1, 0),
    divorced = ifelse(marital_status == 4, 1, 0),
    separated = ifelse(marital_status == 3, 1, 0),
    single = ifelse(marital_status %in% c(0, 1), 1, 0)
  ) %>% 
  group_by(id) %>% 
  #calculate the total number of months of marriage per respondent
  mutate(
    months_married = sum(married, na.rm = TRUE)
  ) %>% 
  ungroup() %>% 
  #only keep respondents who were married for at least 1 month 
  filter(months_married > 0) %>% 
  select(!marital_status)

#data quality checks
#for each respondent, we should have 156 rows (13 years * 12 months) 
#number_of_timepoints <- ncol(NLSY97 %>% select(contains("MAR")))
#number_of_respondents <- length(unique(marital$id)) #in paper: 2,848 (Table 2)
#nrow(marital) == number_of_respondents * number_of_timepoints #TRUE
#"a third of the full sample has ever married by 2009:"
#number_of_respondents / nrow(NLSY97) #35 percent
#mean number of months married in paper is 47.86
#mean(marital$months_married) #48.73 months
#for those who have ever been married, marriage length is an average of 4 years
#mean(marital$months_married) /12 #4.06 years
#all data quality checks passed
```

## Create final, merged dataset

```{r merge}
#keep only the marriage data for respondents who were ever arrested
marital <- marital %>% 
  filter(id %in% arrest$id)

#keep only the arrest data for respondents who were ever married
arrest <- arrest %>% 
  filter(id %in% marital$id)

#keep only age data for respondent who were ever arrested and ever married
all_ages <- all_ages %>% 
  filter(id %in% arrest$id)
 
#create final dataframe (long) with marriage data, arrest data, and covariates
df_long <- left_join(marital, df, by = "id") %>% 
  left_join(arrest, by = c("id", "time")) %>% 
  mutate(
    id = factor(id),
    age = all_ages$age,
    age_sq = age^2
  ) %>%
  group_by(id) %>%
  mutate(
    #mean relationship state for each individual (between-individual control)
    married_mean = mean(married, na.rm = TRUE),
    divorced_mean = mean(divorced, na.rm = TRUE),
    separated_mean = mean(separated, na.rm = TRUE),
    single_mean = mean(single, na.rm = TRUE)
  )

#data quality check: n=813 in paper
#nlevels(df_long$id) #n=882
```


## Statistical model

Information obtained from paper:

- Generalized Mixed Model (binomial family with logit link)
- Estimates obtained using penalized quasi-likelihood estimation (PQL)
- Dependent variable: arrest status (log odds of arrest)
- Time-varying covariates: relationship status (single, marriage, legal separation, divorce), age, and age squared
- Time-invariant covariates: gender, race, household size, family structure, disadvantage, parental education, at-risk, incarceration history
- Level-1 ('within') variables: relationship state, age, and age$^2$
- Level-2 ('between') variables: aggregate relationship state, controls


```{r model, message=FALSE}
library(MASS)
longest_marriage <- max(df_long$months_married)

#Function to run the statistical model for a given number of months married
run_model <- function(months = 1:longest_marriage) {
  #run a generalized mixed model (binomial with logit link)
  glmmPQL( #using penalized quasi-likelihood (PQL) estimation
    #regress arrest on within-individual and between-individual predictors
  fixed = arrest ~ 
    #within-individual, time-varying predictors
    divorced + separated + single + age + age_sq +
    #between-individual, time-stable predictors
    divorced_mean + separated_mean + single_mean + 
    male + black + household_size + family_structure + disadvantage + 
    parental_education + at_risk + incarcerated, 
  random = ~ 1 | id, #each individual is a cluster with 156 timepoints
  family = binomial, #we're predicting the log odds of arrest
  data = df_long %>% filter(months_married %in% months),
  verbose = FALSE
  )
}
```

\newpage

## Claim evaluations

Bushel claims obtained from the OSF [here](https://osf.io/5xqeh/?view_only=0498fe041e4c4753b4126ac28e9960a8). 

::: {.panel-tabset}

## Claim 1


**Coded claim text (original paper):** Divorce has a significant
    detrimental effect on offending and increases the likelihood of
    arrest by 52 percent (table 3, model 2). Stated simply, an
    individual is more criminally active when divorced compared with
    when he or she is married, taking into account periods of legal
    separation and time single.  
    [Table 3, Within Individual: Divorce,
    Model 2, Coeff. = .418, SE = .093, Sig. = p &lt; .001]

**Reproduction data source(s)**:  
[https://www.nlsinfo.org/investigator/pages/search#](https://www.nlsinfo.org/investigator/pages/search#)  

**Description of reproduction data:**  
Original data obtained from the US Bureau of Labor Statistics. NLSY97 (National Longitudinal Study of Youth 1997) downloaded from the NLS investigator on March 1, 2022.
    
    
**Reproduction criteria**

|Criterion          |Original value|Precise reproduction|Approximate reproduction|Non-reproduction|
|------|------|------|--------|--------|
Sample size         |n = 813|n = 813|691 $\le$ n $\le$ 935|n < 691 or n > 935|
Coefficient   |$\beta$ = .418|$\beta$ = .418|.355 $\le \beta \le$ .481|$\beta$ < .355 or $\beta$ > .481|
p-value       |p < .001|p < .001|.001 $\le$ p $\le$ .051|p > .051            |

NB: The sample size and focal coefficient are considered "approximately reproduced" when the reproduction result is within 15% of the original result. For *p*-values, the result is considered "approximately reproduced" when it is within 0.05 points of the original. 

**Analyst success criteria:** The claim will be considered reproducible if a generalized mixed model predicting the log odds of arrest meets the approximate reproduction criteria for the sample size, p-value, and the focal coefficient of divorce. 

**Reproduction results**

```{r g82z5k, cache=TRUE}
claim1 <- run_model() #run model on entire sample
#check fixed effect of divorce
summary(claim1)$tTable["divorced", c("Value", "p-value")]
claim1$dims$ngrps[1] #check sample size
```

**Reproduction outcome:** Based on the reproduction criteria, the claim did reproduce.


## Claim 2: Model 1

**Coded claim text (original paper):** For short marriages, those lasting
    23 months or less (models 1 and 2), neither getting divorced nor
    legally separating statistically alters one’s risk of arrest
    compared with when they were married. Conversely, for those in
    marriages lasting 24 months or longer, divorce is associated with a
    significant increase in the likelihood of arrest.  
    [Table 5. Effect
    of Marital Dissolution on the Probability of Arrest by Marriage
    Length, Within Individual: Divorce, Model 1 Coeff. = .113, SE =
    .228]
    
    
**Reproduction data source(s)**:  
[https://www.nlsinfo.org/investigator/pages/search#](https://www.nlsinfo.org/investigator/pages/search#)  

**Description of reproduction data:**  
Original data obtained from the US Bureau of Labor Statistics. NLSY97 (National Longitudinal Study of Youth 1997) downloaded from the NLS investigator on March 1, 2022.

**Reproduction criteria**

|Criterion          |Original value|Precise reproduction|Approximate reproduction|Non-reproduction|
|------|------|------|--------|--------|
Sample size|n = 121|n = 121|103 $\le$ n $\le$ 139|n < 103 or n > 139|
p-value|p > .05|p > .05|0 < p $\le$ .05|p = 0            |

NB: The sample size and focal coefficient are considered "approximately reproduced" when the reproduction result is within 15% of the original result. For *p*-values, the result is considered "approximately reproduced" when it is within 0.05 points of the original. 


**Analyst success criteria:** The claim will be considered reproducible if a generalized mixed model predicting the log odds of arrest meets the approximate reproduction criteria for the sample size and the precise reproduction criteria for the p-value (p > .05).

**Reproduction results**

```{r model1, cache=TRUE}
#Model 1
model1 <- run_model(months = 1:11) 
summary(model1)$tTable["divorced", "p-value"] #check p-value model 1
model1$dims$ngrps[1] #check sample size model 1
```

**Reproduction outcome:** Based on the reproduction criteria, the claim did reproduce.


## Claim 2: Model 2

**Coded claim text (original paper):** For short marriages, those lasting
    23 months or less (models 1 and 2), neither getting divorced nor
    legally separating statistically alters one’s risk of arrest
    compared with when they were married. Conversely, for those in
    marriages lasting 24 months or longer, divorce is associated with a
    significant increase in the likelihood of arrest.  
    [Table 5. Effect
    of Marital Dissolution on the Probability of Arrest by Marriage
    Length, Within Individual: Divorce, Model 2 Coeff. = -.442, SE = .396]
    
**Reproduction data source(s)**:  
[https://www.nlsinfo.org/investigator/pages/search#](https://www.nlsinfo.org/investigator/pages/search#)  

**Description of reproduction data:**  
Original data obtained from the US Bureau of Labor Statistics. NLSY97 (National Longitudinal Study of Youth 1997) downloaded from the NLS investigator on March 1, 2022.

**Reproduction criteria**

|Criterion          |Original value|Precise reproduction|Approximate reproduction|Non-reproduction|
|------|------|------|--------|--------|
Sample size|n = 120|n = 120|102 $\le$ n $\le$ 138|n < 102 or n > 138|
p-value|p > .05|p > .05|0 < p $\le$ .05|p = 0            |

NB: The sample size and focal coefficient are considered "approximately reproduced" when the reproduction result is within 15% of the original result. For *p*-values, the result is considered "approximately reproduced" when it is within 0.05 points of the original. 

**Analyst success criteria:** The claim will be considered reproducible if a generalized mixed model predicting the log odds of arrest meets the approximate reproduction criteria for the sample size and the precise reproduction criteria for the p-value (p > .05).

**Reproduction results**

```{r model2, cache = TRUE}
#Model 2
model2 <- run_model(months = 12:23) 
summary(model2)$tTable["divorced", "p-value"] #check p-value model 2
model2$dims$ngrps[1] #check sample size model 2
```

**Reproduction outcome:** Based on the reproduction criteria, the claim did reproduce.

## Claim 2: Model 3

**Coded claim text (original paper):** For short marriages, those lasting
    23 months or less (models 1 and 2), neither getting divorced nor
    legally separating statistically alters one’s risk of arrest
    compared with when they were married. Conversely, for those in
    marriages lasting 24 months or longer, divorce is associated with a
    significant increase in the likelihood of arrest.  
    [Table 5. Effect
    of Marital Dissolution on the Probability of Arrest by Marriage
    Length, Within Individual: Divorce, Model 3 Coeff. = .465, SE =
    .151, Sig. = p &lt; .01]
    
**Reproduction data source(s)**:  
[https://www.nlsinfo.org/investigator/pages/search#](https://www.nlsinfo.org/investigator/pages/search#)  

**Description of reproduction data:**  
Original data obtained from the US Bureau of Labor Statistics. NLSY97 (National Longitudinal Study of Youth 1997) downloaded from the NLS investigator on March 1, 2022.

**Reproduction criteria**

|Criterion          |Original value|Precise reproduction|Approximate reproduction|Non-reproduction|
|------|------|------|--------|--------|
Sample size|n = 142|n = 142|121 $\le$ n $\le$ 163|n < 121 or n > 163|
Coefficient|$\beta$ = .465|$\beta$ = .465|.395 $\le \beta \le$ .535|$\beta$ < .395 or $\beta$ > .535|

NB: The sample size and focal coefficient are considered "approximately reproduced" when the reproduction result is within 15% of the original result. For *p*-values, the result is considered "approximately reproduced" when it is within 0.05 points of the original. 

**Analyst success criteria:** The claim will be considered reproducible if a generalized mixed model predicting the log odds of arrest meets the approximate reproduction criteria for the sample size and the approximate reproduction criteria for the coefficient of divorce.

**Reproduction results**

```{r model3, cache = TRUE}
#Model 3
model3 <- run_model(months = 24:35) 
summary(model3)$tTable["divorced", "Value"] #check estimate model 3
model3$dims$ngrps[1] #check sample size model 3
```

**Reproduction outcome:** Based on the reproduction criteria, the claim did reproduce.

## Claim 2: Model 4

**Coded claim text (original paper):** For short marriages, those lasting
    23 months or less (models 1 and 2), neither getting divorced nor
    legally separating statistically alters one’s risk of arrest
    compared with when they were married. Conversely, for those in
    marriages lasting 24 months or longer, divorce is associated with a
    significant increase in the likelihood of arrest.  
    [Table 5. Effect
    of Marital Dissolution on the Probability of Arrest by Marriage
    Length, Within Individual: Divorce, Model 4 Coeff. = .522, SE = .107, Sig. = p
    &lt; .001]
    
**Reproduction data source(s)**:  
[https://www.nlsinfo.org/investigator/pages/search#](https://www.nlsinfo.org/investigator/pages/search#)  

**Description of reproduction data:**  
Original data obtained from the US Bureau of Labor Statistics. NLSY97 (National Longitudinal Study of Youth 1997) downloaded from the NLS investigator on March 1, 2022.

**Reproduction criteria**

|Criterion          |Original value|Precise reproduction|Approximate reproduction|Non-reproduction|
|------|------|------|--------|--------|
Sample size|n = 430|n = 430|366 $\le$ n $\le$ 495|n < 366 or n > 495|
Coefficient|$\beta$ = .522|$\beta$ = .522|.444 $\le \beta \le$ .600|$\beta$ < .444 or $\beta$ > .600|
p-value|p < .001|p < .001|.001 $\le$ p $\le$ .051|p > .051|

NB: The sample size and focal coefficient are considered "approximately reproduced" when the reproduction result is within 15% of the original result. For *p*-values, the result is considered "approximately reproduced" when it is within 0.05 points of the original. 

**Analyst success criteria:** The claim will be considered reproducible if a generalized mixed model predicting the log odds of arrest meets the approximate reproduction criteria for the sample size, the coefficient of divorce, and the p-value.

**Reproduction results**

```{r model4, cache=TRUE}
#Model 4
model4 <- run_model(months = 36:longest_marriage) 
summary(model4)$tTable["divorced", c("Value", "p-value")] #check estimate + p-value model 4
model4$dims$ngrps[1] #check sample size model 4
```

**Reproduction outcome:** Based on the reproduction criteria, the claim did reproduce.



::: 

## Description of materials provided    

All materials on [the OSF project](https://osf.io/3j7gq/?view_only=b0e33287a13146d1a2d9c4ad938be094) may be shared publicly. Overview and description of the materials:

- `BERSANI_Criminology_2013_zmYY_bushel_claims.md`  
An overview of all claims extracted from the paper
- `BERSANI_Criminology_2013_zmYY_2w9mo_SDR_PREREGISTRATION.xlsx`   
Excel workbook with the preregistered reproduction criteria for all claims part of the reproduction attempt
- `NLSY97.dat`   
Data file (National Longitudinal Study of Youth 1997) downloaded from the NLS investigator
- `codebook.txt`  
Codebook (data dictionary) accompanying `NLSY97.dat`
- `helper-functions.R`  
Helper functions accompanying `NLSY97.dat`
- `Analysis-script.Rmd`  
Preregistered analysis script (on randomly shuffled, blinded data)
- `Analysis-script.html`   
Knitted output of the preregistered analysis script (on randomly shuffled, blinded data)
- `Transparency Trail.Rmd`   
Transparency trail of the reproduction attempt. This includes the analysis pipeline and full results/output of the analysis script (applied to the unblinded data)
- `Transparency-Trail.pdf`   
Knitted output of the transparency trail of the reproduction attempt. This includes the analysis pipeline and full results/output of the analysis script (applied to the unblinded data)


## References

Bersani, B. E.; Doherty, E. (2013). When the ties that bind unwind: Examining the enduring and situational processes of change behind the marriage effect. *Criminology, 51*(2), 399. <https://doi.org/10.1111/1745-9125.12008>

Bureau of Labor Statistics, U.S. Department of Labor. National Longitudinal Survey of Youth 1997 cohort, 1997-2009. Produced and distributed by the Center for Human Resource Research (CHRR), The Ohio State University. Columbus, OH: 2019.